A planetary collision afterglow and transit of the resultant debris cloud

### 2023-10-10 src/tex/ms.tex
[showyourwork] overleaf sync

% this is explained at https://github.com/tectonic-typesetting/tectonic/issues/70#issuecomment-974462614
\PassOptionsToClass{pdflatex}{sn-jnl}
\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals

\usepackage{graphicx}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{xcolor}%
\usepackage{manyfoot}%

@@

\newcommand*\degr{\ensuremath{^\circ}}
\newcommand*\arcsec{\ensuremath{^{\prime\prime}}}

@@

\affil[14]{\orgdiv{School of Physics, H H Wills Physics Laboratory}, \orgname{University of Bristol}, \orgaddress{\street{Tyndall Avenue}, \city{Bristol}, \postcode{BS8 1TL}, \country{USA}}}

@@

\abstract{
Planets grow in rotating disks of dust and gas around forming stars, some of which can subsequently collide in giant impacts after the gas component is removed from the disk \cite{Williams11,Wyatt15,Hughes18}.
%
Monitoring programs with the warm Spitzer mission have recorded significant and rapid changes in mid-infrared output for several stars, interpreted as variations in the surface area of warm dusty material ejected by planetary-scale collisions and heated by the central star: e.g., NGC 2354–ID8 \cite{2014Sci...345.1032M,Su19}, HD 166191 \cite{Su22} and V844 Persei \cite{2021ApJ...918...71R}.
%
%However, observational constraints on the aftermath and frequency of such impacts are difficult to obtain.
%
Here we report combined observations of the young ($\sim$~300~Myr), solar-like star \asas{}: an infrared brightening consistent with a blackbody temperature of 1000 K and a luminosity of 4\% of that of the star lasting for about 1000\,days, partially overlapping in time with a complex and deep wavelength-dependent optical eclipse that lasted for about 500 days.
%
The optical eclipse started 2.5\,years after the infrared brightening, implying an orbital period of at least that duration.
%
These observations are consistent with a collision between two exoplanets of several to tens of Earth masses at 2-16\,au from the central star. 
%
Such an impact produces a hot, highly-extended post-impact remnant with sufficient luminosity to explain the infrared observations.
%
Transit of the impact debris, sheared by orbital motion into a long cloud, causes the subsequent complex eclipse of the host star.
}

@@

The otherwise unremarkable star 2MASS J08152329-3859234 underwent a sudden optical dimming event in December 2021 \cite{RizzoSmith21,RizzoSmith22} and was assigned the identifier \asas{} by the ASAS-SN survey \citep{shappee_man_2014,kochanek_all-sky_2017}.
%
Here, we combine both optical (from the Las Cumbres Observatory Global Telescope, LCOGT) and infrared (from the WISE satellite) observations of \asas~for the years before and after this dimming event (Figure~\ref{fig:wisephot}). 
%
Optical multiband photometry shows a wavelength dependent depth consistent with extinction by sub-micron particles.
%
About 900 days prior to the optical dimming event, the \asas~system showed a significant brightening in the infrared, of 0.4 magnitudes at 3.8 microns ($W1$) and 0.8 magnitudes at 4.5 microns ($W2$).
%
Before this time the IR brightness was consistent with being purely stellar.
%
The increased IR fluxes indicate that in addition to the quiescent stellar flux, there was new emission at a temperature of approximately 1000\,K.
%
Such a remarkable combination of observations, particularly the 2.5\,year delay between the IR and optical variation, requires an explanation.

\begin{figure}
\begin{centering}
\includegraphics[width=\textwidth]{figures/all_phot_nature_delta_flux.pdf}
%\includegraphics[width=1.0\textwidth]{figures/master_lightcurve_nature_simple.pdf}
\caption{Optical and infrared photometry of ASASSN-21qj.
%
Panel {\bf a} is normalised optical photometry from ASAS-SN in $V$-band and $g'$ band.
%
Panel {\bf b} shows the fractional flux increase in brightness of ASASSN-21qj in both the $W1$ and $W2$ bands, where a value of 1.0 represents the stellar contribution alone.
%
Panel {\bf c} shows the calculated $NEOWISE$ color temperature estimated from the photometry of the two bands.
%
The color temperature is plotted as zero when there is no IR excess, and is consistent with a temperature of 1000\,K while the excess is present.
%
Error bars are shown at $1\sigma$ confidence.
}
\label{fig:wisephot}
\script{plot_all_phot_nature_delta_flux.py}
%\script{plot_master_lightcurve_nature_simple.py}
\end{centering}
\end{figure}

@@

The optical and infrared light curves (Figure \ref{fig:wisephot}) provide key constraints on any proposed scenario.
%
The post-brightening IR fluxes in the $W1$ and $W2$ WISE passbands are consistent with emission at a blackbody temperature of $1000 \pm 100$\,K and this temperature is sustained, within error, for the remainder of our observation window, despite a decline in the total flux.
%
For an emitter located at the distance of ASASSN-21qj from Earth, and the observed  maximum luminosity of approximately 0.04\,$L_\star$, this implies an emitting area of 0.01\,au$^2$, equivalent to an object with a radius of 7\,$R_\odot$, or $750$\,$R_{\rm Earth}$.
%
If this emission was from material -- e.g., dust -- passively heated by proximity to the star, then that material must have been generated and remained within about 0.1\,au to produce the observed temperature.

\begin{figure*}
\begin{centering}
\includegraphics[width=1.0\textwidth]{figures/master_lightcurve_nature_simple.pdf}
      \caption{The light curve of ASASSN-21qj from several different photometric surveys and the derived transverse velocities.
      %
      Panel {\bf a} shows that the eclipse depth is deeper for shorter wavelengths, indicating that the transiting material is dominated by sub-micron sized grains.
      %
      Panel {\bf b} shows the  transverse velocities derived from the light curve gradients.
      %
      These are lower limits to the true velocity, and thus imply that the transiting material is closer to the star than 16\,au.
      %
      Error bars are shown at $1\sigma$ confidence.
      }
        \label{fig:eclipse_overview}
\script{plot_master_lightcurve_nature_simple.py}
\end{centering}
\end{figure*}

@@

Instead, we hypothesize that are we are observing the aftermath of a single collision between super-Earths or mini Neptunes -- a so-called giant impact -- between 2 and 16\,au from the star. 
%
These distances are determined, respectively, by the delay between the IR brightening and the optical eclipse (Fig. \ref{fig:wisephot}), and by gradients in the optical light curve (Fig. \ref{fig:eclipse_overview}).
%
In contrast to other extreme debris disk events where the star heats the dust, we propose that the infrared emission is directly from the post-impact body \cite{Lock2017,2009ApJ...704..770M}, and that impact debris produced the optical transit.
%
Giant impacts are a common occurrence in planet formation \cite{Schlichting2018a,DAngelo2018} and also occur during instabilities in older systems \cite{Kaib2016}; this would explain the observations with a single event of a type that are expected for systems with ages like \asas{}.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=\columnwidth]{figures/Size_cooling_combined_figure.pdf}
\caption{The size and temporal evolution of a post-impact body.
%
A post-impact body of a few to tens of Earth masses could be large enough to explain the observed increase in thermal emission from ASASSN-21qj and the subsequent infrared fluxes.
%
Panel {\bf a} shows the Hill radii for bodies at different semi-major axis (solid colored lines) and the Bondi radii (dotted lines) assuming different compositions for the vapor of the post-impact body.
%
The horizontal black dashed line shows the radii required to explain the observed flux at the inferred emission temperature of 1000~K. 
%
The change in excess flux due to a post-impact body with time for a simple model of cooling of a post-impact body for different power-law surface density profiles is in panel {\bf b}, eqn~\ref{eqn:sigma} and for different mass bodies in panel {\bf c}.
%
Each example had an initial radius of 7\,$R_\odot$, a radiative temperature of 1000~K, and a corresponding initial flux of 0.04\,$L_\star$, as estimated for the observed infrared emittor.
%
The solid lines are profiles that can have non-zero initial density at the initial emitting radius, and the dashed lines are for ones that are forced to have zero density at the initial emitting radius.
%
When not stated, the mass of the post-impact body was 50~$M_{\rm Earth}$, the power law exponent is $-2$, and the size of the central region was 10~$R_{\rm Earth}$.
}
\label{fig:Hill_Bondi_R}
\script{plot_hill_bondi_radii_three_panels.py}

@@

Giant impacts produce significant amounts of debris, typically around 1\% of the colliding mass, that is injected into orbit around the host star\cite{Canup2001,Lock18}.
%
Impact ejecta has a wide range of sizes, from sub-micron dust to planetesimals of tens to hundreds of kilometers, and often contains the most highly heated material \cite{Benz2008_Mercury_book,Leinhardt2015,Carter2020a}.
%
For sufficiently high impact velocities ($>1$~km~s$^{-1}$ for water ice, and $>8$~km~s$^{-1}$ for forsterite) a substantial fraction of this material is vaporised \cite{Stewart2008,Davies2020,Carter2020a}.
%
Shearing of droplets and cooling and condensation of vapor produces a population of small dust grains and solid spherules.
%
The size distribution of this fraction of the debris is uncertain, due to difficulties in modelling condensate nucleation and breakup, but previous work suggests that debris could range in size from sub-micron to decimeters \cite{Benz2008_Mercury_book,Johnson2015}.
%
The wavelength-dependent eclipse suggests that the optical depth of the transiting dust cloud is dominated by sub-micron grains, consistent with these previous estimates.
%
While the fading of the excess infrared flux occurs within 100 days of the start of the optical transit, we consider the timing to be coincidental because there is no clear correspondence between the light curves, for example no  change in the (six month cadence) IR flux when the star dims in the optical wavelengths just before MJD 59500.

@@

The radiative flux from post-impact bodies has not been explored in depth.
%
Computational resource limitations make resolving the low-density outer regions and photosphere of post-impact bodies extremely challenging.
%
However, preliminary simulations of impacts between super-Earth and mini-Neptunes place an approximate lower limit on the extent of post-impact bodies and show that post-impact bodies can extend to hundreds of Earth radii.
%
Such an object radiating at $\sim$1000~K would produce a flux comparable to the 0.04~$L_*$ inferred from our observations.

Independent of impact simulations, there are fundamental limits on the size of post-impact bodies from the Hill and Bondi radii.
%
Any post-impact structure must lie within the Hill sphere, the distance within which the gravity of an object dominates over that of the star.
%
Furthermore, only vapor within the Bondi radius would be bound to the post-impact body.
%
Figure~\ref{fig:Hill_Bondi_R}A shows the Hill radii at different distances from ASASSN-21qj (solid lines) and the Bondi radius for example gas species (dotted lines).
%
Beyond 2.4\,au, the Hill radii of greater than Earth-mass bodies are large enough to accommodate a post-impact body capable of producing the required IR flux ($\sim7R_*$, dotted line). 
%
Heavier gases (H$_2$O, SiO, and SiO$_2$) are also bound to bodies of more than a few Earth masses.
%
A post-impact body of a few Earth masses can hence theoretically produce the observed IR emission, with a photosphere that is dust and/or vapor.

@@

A key line of evidence for direct detection of a post-impact body is the constant emission temperature.
%
It is argued \cite{Lock18} that rocky post-impact bodies become optically thin at low pressures where radiative loss drives rapid condensation of the rock vapor.
%
The emission temperature is then set by the liquid-vapor phase boundary and is constant until the post-impact body almost fully condenses \cite{Lock18,Caracas2023}.
%
The dew point (onset of condensation) and bubble point (onset of substantial vaporization) of material with bulk silicate Earth composition are similar \cite[$\sim2300$~K, within $\sim 100$~K;][]{Lock18,Fegley2023_BSE_cond} at the low pressures of the photosphere of a post-impact body.
%
However, even small amounts of water (10$^{-3}$ mole fraction) may lower the bubble and dew points for silicates by $\sim$100~K \cite[][]{Fegley2023_BSE_cond,Lock18}. 
%
The emission temperature of bodies produced by collisions of proto-planets composed of rock and ices/volatiles could be buffered at $\sim$1000~K during early evolution.

@@

The temporal variation of flux from a post-impact body is controlled by evolution of the bodies' size which is governed by a number of competing factors \cite{Lock2018moon,Lock2020}, including radiative energy loss, viscous spreading, and mass and angular momentum transfer by condensates.
%
For silicate-dominated post-impact bodies, the high emission temperature means that radiative cooling dominates, and the post-impact body contracts rapidly, fully condensing over years to thousands of years \cite{Lock2018moon,Lock2020}.
%
The body we observed has a much lower emission temperature, and contracted substantially more slowly.
%
Figure~\ref{fig:Hill_Bondi_R}~B and C show evolution of conceptual post-impact bodies in the limiting case that radiative cooling dominates.
%
If sufficient mass is injected into the outer regions (i.e., less negative power law exponents) the observed flux can remain constant for an initial period and then decays over the order of hundreds of days, in agreement with infrared observations.
%
Further work is required to understand the structure and evolution of bodies produced by impacts between super-Earths and mini-Neptunes of different compositions and identify temporal flux variations consistent with our observations.

@@

\renewcommand\figurename{Extended Data Fig.}% defined as per springer style 
\renewcommand\tablename{Extended Data Table}%

@@

The stellar properties of \asas\ (Gaia DR3 5539970601632026752 = 2MASS~J08152329-3859234) are listed in Extended Data Table~\ref{tab:Stellarprop}, showing that \asas\ is consistent with being a G2 type dwarf star.
%
Where necessary we assume a stellar mass equal to the Sun.
%
\asas\  has a neighbor (Gaia DR3 5539970597334497024 = 2MASS~J08152298-3859244) which is a visual double.
%
Based on the Gaia EDR3 mean ICRS position for epoch 2016.0, the visual companion lies at a separation $\rho$ = $3738.243\pm0.062$ mas and at position angle $\theta$ = $249^{\circ}.977$.
%
Their parallaxes ($\varpi$ = $1.7631\pm0.0112$ mas vs. $1.4711\pm0.0523$ mas) differ by 5.5$\sigma$ and proper motions ($\mu_{\alpha} = -9.692\pm0.012$, $\mu_{\delta} = 7.349\pm0.012$ \masyr\, vs. $\mu_{\alpha} = -0.114\pm0.055$, $\mu_{\delta} = 6.419\pm0.053$ \masyr) differ by a factor of 2.
%
The large differences in distance and proper motions suggests that these stars are not associated.

The stellar photospheric flux was estimated by fitting stellar models to GAIA, APASS, and DENIS and WISE optical/near-IR photometry (the 2MASS $J$ photometry is an upper limit, and $H$ and $K_s$ are flagged as contaminated).
%
Extended Data Figure \ref{fig:sed} shows the resulting models (the dashed line is discussed below).
%
The ALLWISE photometry ($\sim$2010) is consistent with the NEOWISE photometry pre-brightening nearly ten years later.
%

A fit using the method of \cite{2019MNRAS.488.3588Y} with GAIA and DENIS photometry finds that the WISE W1/2 fluxes are about 20\% too high, but better agreement is found with APASS photometry.
%
The difference is explained by the fact that WISE and APASS have lower spatial resolution and include the flux of the $\sim$2\,mag fainter visual double to the West of \asas{} (which is visible in 2MASS).
%
The best fit stellar effective temperature is $5560 \pm 100$\,K.
%
We do not find that reddening is needed for these models; while there is relatively little photometry with which to strongly constrain both $T_{\rm eff}$ and $A_V$, the $T_{\rm eff}$ from GAIA DR3 is $5760 \pm 10$\,K, which suggests that the conclusion of little reddening is valid.
%
Primarily, we conclude that there is no evidence that \asas{} showed evidence for an IR excess before the brightening seen by NEOWISE.

To estimate the IR excess properties we fit the same models, but now including the first five post-brightening NEOWISE data points.
%
We correct for the nearby source by using the GAIA photometry for the star, and the post/pre-brightening difference for the WISE excess flux.
%
This fit yields the fractional luminosity $L_{\rm dust}/L_\star = 0.04 \pm 0.005$ and a dust temperature $950 \pm 30$\,K.
%
To estimate the dust temperature as a function of time we simply subtract the median pre-brightening W1/2 fluxes, with the RMS of these values as the uncertainty (see Figure \ref{fig:wisephot}).
%
The temperature uncertainty increases as the excess fades; the excess flux uncertainty depends on both the observed flux and the stellar flux, and as the excess decreases the stellar uncertainty (which is constant) becomes an increasingly large fraction of the excess.
%
These fluxes therefore exclude the contaminating flux from the nearby object, and are consistent with the SED-derived dust temperature.

Converting dust temperatures to stellocentric radii is uncertain because dust temperature depends on grain size.
%
Typically the radius derived under the assumption of blackbody emission is an underestimate by a factor of up to five \citep{2013MNRAS.428.1263B,2015MNRAS.454.3207P}.
%
Thus, we conclude that the dust location might in the most extreme case be as far as 1\,au, but not sufficiently far to explain the 900\,day delay between the WISE and optical flux variations. which requires a distance of at least 2\,au.

@@

The beginning of the eclipse was announced \cite{RizzoSmith21} by the ASAS-SN survey, which triggered several observing campaigns at optical wavelengths and an ALMA observation at Band 7 (program \texttt{2019.A.00040.S}).
%
The photometric data and passband filters are listed in Extended Data Table~\ref{tab:photometry}.

@@

In this section we consider the light curves; first the implications of the TESS data for the stellar age, and then the implications of the ground-based optical light curves for transverse velocity of the occulting material and dust grain sizes.

\subsection*{Stellar rotation and age}

>>>>

% this is explained at https://github.com/tectonic-typesetting/tectonic/issues/70#issuecomment-974462614
\PassOptionsToClass{pdflatex}{sn-jnl}
\documentclass[sn-nature,oneside]{sn-jnl}% Style for submissions to Nature Portfolio journals
%%%NOTE: The documentclass has been edited

\usepackage{graphicx}%
\usepackage{amsmath,amssymb}%
%\usepackage{xcolor}% % already required in cls
\usepackage{manyfoot}%

@@

\usepackage{siunitx}
\DeclareSIUnit\jansky{\text{Jy}}

\usepackage{glossaries}
\glsdisablehyper    % no hyperlink because we're not printing the glossary
%% ACRONYMS %%
\newacronym{asassn}{ASAS-SN}{All Sky Automated Survey for Supernovae}
\newacronym{asas}{ASAS}{All Sky Automated Survey}
\newacronym{neowise}{NEOWISE}{Near-Earth Object Wide-field Infrared Survey Explorer}
\newacronym{lcogt}{LCOGT}{Las Cumbres Observatory Global Telescope}
\newacronym{atlas}{ATLAS}{Asteroid Terrestrial-impact Last Alert System}
\newacronym{aavso}{AAVSO}{American Association of Variable Star Observers}
\newacronym{banzai}{BANZAI}{Beautiful Algorithms to Normalize Zillions of Astronomical Images}
\newacronym{road}{ROAD}{Remote Observatory Atacama Desert}
\newacronym{ccd}{CCD}{charge-coupled device}
\newacronym{wise}{WISE}{Wide-field Infrared Survey Explorer}
\newacronym{sed}{SED}{spectral energy distribution}
\newacronym{tess}{TESS}{Transiting Exoplanet Survey Satellite}
\newacronym{alma}{ALMA}{Atacama Large Millimeter Array}
\newacronym{mast}{MAST}{Mikulski Archive for Space Telescopes}

\newcommand*\degr{\ensuremath{^\circ}}
\newcommand*\arcsec{\ensuremath{^{\prime\prime}}}   % note, if you use siunitx package these are included

@@

\affil[14]{\orgdiv{School of Physics, H H Wills Physics Laboratory}, \orgname{University of Bristol}, \orgaddress{\street{Tyndall Avenue}, \city{Bristol}, \postcode{BS8 1TL}, \country{UK}}}

@@

\abstract{
Planets grow in rotating disks of dust and gas around forming stars, some of which can subsequently collide in giant impacts after the gas component is removed from the disk \cite{Williams11,Wyatt15,Hughes18}.
%
Monitoring programs with the warm Spitzer mission have recorded significant and rapid changes in mid-infrared output for several stars, interpreted as variations in the surface area of warm dusty material ejected by planetary-scale collisions and heated by the central star: e.g., NGC 2354–ID8 \cite{2014Sci...345.1032M,Su19}, HD 166191 \cite{Su22} and V844 Persei \cite{2021ApJ...918...71R}.
%
%However, observational constraints on the aftermath and frequency of such impacts are difficult to obtain.
%
Here we report combined observations of the young ($\sim$~300~Myr), solar-like star \asas{}: an infrared brightening consistent with a blackbody temperature of 1000 K and a luminosity that is 4\% that of the star lasting for about 1000\,days, partially overlapping in time with a complex and deep wavelength-dependent optical eclipse that lasted for about 500 days.
%
The optical eclipse started 2.5\,years after the infrared brightening, implying an orbital period of at least that duration.
%
These observations are consistent with a collision between two exoplanets of several to tens of Earth masses at 2-16\,au from the central star. 
%
Such an impact produces a hot, highly-extended post-impact remnant with sufficient luminosity to explain the infrared observations.
%
Transit of the impact debris, sheared by orbital motion into a long cloud, causes the subsequent complex eclipse of the host star.
}

@@

The otherwise unremarkable star 2MASS J08152329-3859234 underwent a sudden optical dimming event in December 2021 \cite{RizzoSmith21,RizzoSmith22} and was assigned the identifier \asas{} by the \gls{asassn} \citep{shappee_man_2014,kochanek_all-sky_2017}.
%
Here, we combine both optical (from the \gls{lcogt}) and infrared (from the \gls{wise} satellite) observations of \asas~for the years before and after this dimming event (Figure~\ref{fig:wisephot}). 
%
Optical multiband photometry shows a wavelength dependent depth consistent with extinction by sub-micron particles.
%
About 900 days prior to the optical dimming event, the \asas~system showed a notable brightening in the infrared, of 0.4 magnitudes at 3.8 microns ($W1$) and 0.8 magnitudes at 4.5 microns ($W2$).
%
Before this time the infrared brightness was consistent with being purely stellar.
%
The increased infrared fluxes indicate that in addition to the quiescent stellar flux, there was new emission at a temperature of approximately 1000\,K.
%
Such a notable combination of observations, particularly the 2.5\,year delay between the infrared and optical variation, requires an explanation.

\begin{figure}
\begin{centering}
\includegraphics[width=\textwidth]{figures/all_phot_nature_delta_flux.pdf}
%\includegraphics[width=1.0\textwidth]{figures/master_lightcurve_nature_simple.pdf}
\caption{\textbf{Optical and infrared photometry of ASASSN-21qj.}
%
{\bf a}, Normalized optical photometry from \gls{asassn} in the $V$-band and the $g'$ band.
%
{\bf b},Fractional flux increase in brightness of ASASSN-21qj in both the $W1$ and $W2$ bands, for which a value of 1.0 represents the stellar contribution alone.
%
{\bf c},Calculated NEOWISE color temperature estimated from the photometry of the two bands.
%
The color temperature is plotted as zero when there is no infrared excess and is consistent with a temperature of 1000\,K while the excess is present.
%
Error bars are shown at $1\sigma$ confidence.
}
\label{fig:wisephot}
\script{plot_all_phot_nature_delta_flux.py}
%\script{plot_master_lightcurve_nature_simple.py}
\end{centering}
\end{figure}

@@

The optical and infrared light curves (Figure \ref{fig:wisephot}) provide key constraints on any proposed scenario.
%
The post-brightening infrared fluxes in the $W1$ and $W2$ \gls{wise} passbands are consistent with emission at a blackbody temperature of $1000 \pm 100$\,K and this temperature is sustained, within error, for the remainder of our observation window, despite a decline in the total flux.
%
For an emitter located at the distance of ASASSN-21qj from Earth, and the observed  maximum luminosity of approximately 0.04\,$L_\star$, this implies an emitting area of 0.01\,au$^2$, equivalent to an object with a radius of 7\,$R_\odot$, or $750$\,$R_{\rm Earth}$.
%
If this emission was from material -- such as dust -- passively heated by proximity to the star, then that material must have been generated and remained within about 0.1\,au to produce the observed temperature.

\begin{figure*}
\begin{centering}
\includegraphics[width=1.0\textwidth]{figures/master_lightcurve_nature_simple.pdf}
      \caption{\textbf{The light curve of ASASSN-21qj from several different photometric surveys and the derived transverse velocities.}
      %
      {\bf a},The eclipse depth is deeper for shorter wavelengths, indicating that the transiting material is dominated by sub-micron sized grains.
      %
      {\bf b},Transverse velocities derived from the light-curve gradients.
      %
      These are lower limits to the true velocity, and thus imply that the transiting material is closer to the star than 16\,au.
      %
      Error bars are shown at $1\sigma$ confidence.
      }
        \label{fig:eclipse_overview}
\script{plot_master_lightcurve_nature_simple.py}
\end{centering}
\end{figure*}

@@

Instead, we suggest that we are observing the aftermath of a single collision between super-Earths or mini Neptunes -- a so-called giant impact -- between 2 and 16\,au from the star. 
%
These distances are determined, respectively, by the delay between the infrared brightening and the optical eclipse (Figure \ref{fig:wisephot}), and by gradients in the optical light curve (Figure \ref{fig:eclipse_overview}).
%
In contrast to other extreme debris disk events where the star heats the dust, we propose that the infrared emission is directly from the post-impact body \cite{Lock2017,2009ApJ...704..770M}, and that impact debris produced the optical transit.
%
Giant impacts are a common occurrence in planet formation \cite{Schlichting2018a,DAngelo2018} and also occur during instabilities in older systems \cite{Kaib2016}; this would explain the observations with a single event of a type that are expected for systems with ages like \asas{}.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=\columnwidth]{figures/Size_cooling_combined_figure.pdf}
\caption{\textbf{The size and temporal evolution of a post-impact body.}
%
A post-impact body of a few to tens of Earth masses could be large enough to explain the observed increase in thermal emission from ASASSN-21qj and the subsequent infrared fluxes.
%
{\bf a},The Hill radii for bodies at different semi-major axes (solid colored lines) and the Bondi radii (dashed lines) assuming different compositions for the vapor of the post-impact body.
%
The horizontal black dashed line shows the radius required to explain the observed flux at the inferred emission temperature of 1000~K. 
%
{\bf b,c},The change in excess flux owing to a post-impact body with time for a simple model of cooling of a post-impact body for different power-law surface density profiles ({\bf b}; equation~\ref{eqn:sigma}) and for different mass bodies ({\bf c}).
%
Each example had an initial radius of 7\,$R_\odot$, a radiative temperature of 1000~K, and a corresponding initial flux of 0.04\,$L_\star$, as estimated for the observed infrared emitter.
%
The solid lines are profiles that can have non-zero initial density at the initial emitting radius, and the dashed lines are for ones that are forced to have zero density at the initial emitting radius.
%
When not stated, the mass of the post-impact body was 50~$M_{\rm Earth}$, the power law exponent is $-2$, and the size of the central region was 10~$R_{\rm Earth}$.
}
\label{fig:Hill_Bondi_R}
\script{plot_hill_bondi_radii_three_panels.py}

@@

Giant impacts produce substantial amounts of debris, typically around 1\% of the colliding mass, that is injected into orbit around the host star\cite{Canup2001,Lock18}.
%
Impact ejecta have a wide range of sizes, from sub-micron dust to planetesimals of tens to hundreds of kilometers, and often contain the most highly heated material \cite{Benz2008_Mercury_book,Leinhardt2015,Carter2020a}.
%
For sufficiently high impact velocities ($>1$~km~s$^{-1}$ for water ice, and $>8$~km~s$^{-1}$ for forsterite) a substantial fraction of this material is vaporised \cite{Stewart2008,Davies2020,Carter2020a}.
%
Shearing of droplets and cooling and condensation of vapor produces a population of small dust grains and solid spherules.
%
The size distribution of this fraction of the debris is uncertain, owing to difficulties in modelling condensate nucleation and breakup, but previous work suggests that debris could range in size from sub-micron to decimeters \cite{Benz2008_Mercury_book,Johnson2015}.
%
The wavelength-dependent eclipse suggests that the optical depth of the transiting dust cloud is dominated by sub-micron grains, consistent with these previous estimates.
%
Although the fading of the excess infrared flux occurs within 100 days of the start of the optical transit, we consider the timing to be coincidental because there is no clear correspondence between the light curves, for example no  change in the (six month cadence) infrared flux when the star dims in the optical wavelengths just before MJD 59500.

@@

The radiative flux from post-impact bodies has not been explored in depth.
%
Computational resource limitations make resolving the low-density outer regions and photosphere of post-impact bodies extremely challenging.
%
However, preliminary simulations of impacts between super-Earth and mini-Neptunes place an approximate lower limit on the extent of post-impact bodies and show that post-impact bodies can extend to hundreds of Earth radii.
%
Such an object radiating at about 1000~K would produce a flux comparable with the 0.04~$L_\star$ inferred from our observations.

Independent of impact simulations, there are fundamental limits on the size of post-impact bodies from the Hill and Bondi radii.
%
Any post-impact structure must lie within the Hill sphere, the distance within which the gravity of an object dominates over that of the star.
%
Furthermore, only vapor within the Bondi radius would be bound to the post-impact body.
%
Figure~\ref{fig:Hill_Bondi_R}a shows the Hill radii at different distances from ASASSN-21qj (solid lines) and the Bondi radii for example gas species (dotted lines).
%
Beyond 2.4\,au, the Hill radii of greater than Earth-mass bodies are large enough to accommodate a post-impact body capable of producing the required infrared flux (approximately $7R_\star$, dotted line). 
%
Heavier gases (H$_2$O, SiO, and SiO$_2$) are also bound to bodies of more than a few Earth masses.
%
A post-impact body of a few Earth masses can hence theoretically produce the observed infrared emission, with a photosphere that is dust and/or vapor.

@@

A key line of evidence for direct detection of a post-impact body is the constant emission temperature.
%
It is argued \cite{Lock18} that rocky post-impact bodies become optically thin at low pressures where radiative loss drives rapid condensation of the rock vapor.
%
The emission temperature is then set by the liquid-vapor phase boundary and is constant until the post-impact body almost fully condenses \cite{Lock18,Caracas2023}.
%
The dew point (onset of condensation) and bubble point (onset of substantial vaporization) of material with bulk silicate Earth composition are similar (about 2300 K, within roughly 100 K \cite[][]{Lock18,Fegley2023_BSE_cond}) at the low pressures of the photosphere of a post-impact body.
%
However, even small amounts of water (10$^{-3}$ mole fraction) may lower the bubble and dew points for silicates by approximately 100~K \cite[][]{Fegley2023_BSE_cond,Lock18}. 
%
The emission temperature of bodies produced by collisions of proto-planets composed of rock and ices/volatiles could be buffered at about 1000~K during early evolution.

@@

The temporal variation of flux from a post-impact body is controlled by evolution of the bodies' size which is governed by a number of competing factors \cite{Lock18,Lock2020}, including radiative energy loss, viscous spreading, and mass and angular momentum transfer by condensates.
%
For silicate-dominated post-impact bodies, the high emission temperature means that radiative cooling dominates, and the post-impact body contracts rapidly, fully condensing over years to thousands of years \cite{Lock18,Lock2020}.
%
The body we observed has a much lower emission temperature, and contracted much more slowly.
%
Figure~\ref{fig:Hill_Bondi_R}b,c shows the evolution of conceptual post-impact bodies in the limiting case that radiative cooling dominates.
%
If sufficient mass is injected into the outer regions (i.e., less negative power law exponents) the observed flux can remain constant for an initial period and then decays over the order of hundreds of days, in agreement with infrared observations.
%
Further work is required to understand the structure and evolution of bodies produced by impacts between super-Earths and mini-Neptunes of different compositions and identify temporal flux variations consistent with our observations.

@@

\renewcommand\figurename{Extended Data Figure}% defined as per springer style 
\renewcommand\tablename{Extended Data Table}%

@@

The stellar properties of \asas\ (Gaia DR3 5539970601632026752 = 2MASS~J08152329-3859234) are listed in Extended Data Table~\ref{tab:Stellarprop}, showing that \asas\ is consistent with being a G2 type dwarf star.
%
Where necessary we assume a stellar mass equal to the Sun.
%
\asas\  has a neighbor (Gaia DR3 5539970597334497024 = 2MASS~J08152298-3859244) which is a visual double.
%
Based on the Gaia DR3 mean ICRS position for epoch 2016.0, the visual companion lies at a separation $\rho$ = $3738.243\pm0.062$ mas and at position angle $\theta$ = $249^{\circ}.977$.
%
Their parallaxes ($\varpi$ = $1.7631\pm0.0112$ mas vs. $1.4711\pm0.0523$ mas) differ by 5.5$\sigma$ and proper motions ($\mu_{\alpha} = -9.692\pm0.012$, $\mu_{\delta} = 7.349\pm0.012$ \masyr\, vs. $\mu_{\alpha} = -0.114\pm0.055$, $\mu_{\delta} = 6.419\pm0.053$ \masyr) differ by a factor of 2.
%
The large differences in distance and proper motions suggests that these stars are not associated.

The stellar photospheric flux was estimated by fitting stellar models to GAIA, APASS, and DENIS and \gls{wise} optical/near-IR photometry (the 2MASS $J$ photometry is an upper limit, and $H$ and $K_s$ are flagged as contaminated).
%
Extended Data Figure \ref{fig:sed} shows the resulting models (the dashed line is discussed below).
%
The ALLWISE photometry ($\sim$2010) is consistent with the \gls{neowise} photometry pre-brightening nearly ten years later.
%

A fit using the method of \cite{2019MNRAS.488.3588Y} with GAIA and DENIS photometry finds that the \gls{wise} W1/2 fluxes are about 20\% too high, but better agreement is found with APASS photometry.
%
The difference is explained by the fact that \gls{wise} and APASS have lower spatial resolution and include the flux of the approximately 2\,mag fainter visual double to the west of \asas{} (which is visible in 2MASS).
%
The best fit stellar effective temperature is $5560 \pm 100$\,K.
%
We do not find that reddening is needed for these models; while there is relatively little photometry with which to strongly constrain both $T_{\rm eff}$ and $A_V$, the $T_{\rm eff}$ from GAIA DR3 is $5760 \pm 10$\,K, which suggests that the conclusion of little reddening is valid.
%
Primarily, we conclude that there is no indication that \asas{} showed evidence for an infrared excess before the brightening seen by \gls{neowise}.

To estimate the infrared excess properties we fit the same models, but now including the first five post-brightening \gls{neowise} data points.
%
We correct for the nearby source by using the GAIA photometry for the star, and the post/pre-brightening difference for the \gls{wise} excess flux.
%
This fit yields the fractional luminosity $L_{\rm dust}/L_\star = 0.04 \pm 0.005$ and a dust temperature $950 \pm 30$\,K.
%
To estimate the dust temperature as a function of time we simply subtract the median pre-brightening W1/2 fluxes, with the r.m.s. of these values as the uncertainty (see Figure \ref{fig:wisephot}).
%
The temperature uncertainty increases as the excess fades; the excess flux uncertainty depends on both the observed flux and the stellar flux, and as the excess decreases the stellar uncertainty (which is constant) becomes an increasingly large fraction of the excess.
%
These fluxes therefore exclude the contaminating flux from the nearby object, and are consistent with the  \gls{sed}-derived dust temperature.

Converting dust temperatures to stellocentric radii is uncertain because dust temperature depends on grain size.
%
Typically the radius derived under the assumption of blackbody emission is an underestimate by a factor of up to five \citep{2013MNRAS.428.1263B,2015MNRAS.454.3207P}.
%
Thus, we conclude that the dust location might, in the most extreme case, be as far as 1\,au, but not sufficiently far to explain the 900\,day delay between the \gls{wise} and optical flux variations. which requires a distance of at least 2\,au.

@@

The beginning of the eclipse was announced \cite{RizzoSmith21} by the \gls{asassn} survey, which prompted several observing campaigns at optical wavelengths and an \gls{alma} observation at Band 7 (program \texttt{2019.A.00040.S}).
%
The photometric data and passband filters are listed in Extended Data Table~\ref{tab:photometry}.

@@

In this section we consider the light curves; first the implications of the \gls{tess} data for the stellar age, and then the implications of the ground-based optical light curves for transverse velocity of the occulting material and dust grain sizes.

\subsubsection*{Stellar rotation and age}
